---
title: "compare_gistic2_output_nbl"
author: "Ryan Rebernick"
date: "6/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description:



# Load packages
```{r, message=F}

library(dplyr)
library(magrittr)
library(data.table)
library(tidyr)
library(GenomicRanges)

```



# Load data

## nbl cohort gistic
```{r}

ours <- fread('/mctp/share/users/rebernrj/projects/nbl/data/gistic2/all_lesions.conf_99.txt', data.table = F) %>%
  dplyr::filter(`Amplitude Threshold` != 'Actual Copy Change Given') %>%
  dplyr::filter(`Residual q values after removing segments shared with higher peaks` < 0.05)
ours <- ours[,c(1,2,7)]
ours %<>% set_colnames(c('name', 'region', 'residual_q')) 
ours$name.region = paste0(vapply(strsplit(ours$name," "), `[`, 1, FUN.VALUE=character(1)), '.', ours$region)
ours$group <- 'our_nbl_cohort'

```

## depudyt (MYCN = 0)
```{r}

mycn0 <- fread('/mctp/share/users/rebernrj/projects/depuydt_nbl/out/gistic2/MYCN0/all_lesions.conf_90.txt', data.table = F) %>%
  dplyr::filter(`Amplitude Threshold` != 'Actual Copy Change Given') %>%
  dplyr::filter(`Residual q values after removing segments shared with higher peaks` < 0.05)
mycn0 <- mycn0[,c(1,2,7)]
mycn0 %<>% set_colnames(c('name', 'region', 'residual_q')) 
mycn0$name.region = paste0(vapply(strsplit(mycn0$name," "), `[`, 1, FUN.VALUE=character(1)), '.', mycn0$region)
mycn0$group <- 'dep_mycn0'

```

## depudyt (MYCN = 1)
```{r}

mycn1 <- fread('/mctp/share/users/rebernrj/projects/depuydt_nbl/out/gistic2/MYCN1/all_lesions.conf_90.txt', data.table = F) %>%
  dplyr::filter(`Amplitude Threshold` != 'Actual Copy Change Given') %>%
  dplyr::filter(`Residual q values after removing segments shared with higher peaks` < 0.05)
mycn1 <- mycn1[,c(1,2,7)]
mycn1 %<>% set_colnames(c('name', 'region', 'residual_q')) 
mycn1$name.region = paste0(vapply(strsplit(mycn1$name," "), `[`, 1, FUN.VALUE=character(1)), '.', mycn1$region)
mycn1$group <- 'dep_mycn1'

```

## depudyt cases (case = patients that die within 1.5 years)
```{r}

cases <- fread('/mctp/share/users/rebernrj/projects/depuydt_nbl/out/gistic2/cases/all_lesions.conf_90.txt', data.table = F) %>%
  dplyr::filter(`Amplitude Threshold` != 'Actual Copy Change Given') %>%
  dplyr::filter(`Residual q values after removing segments shared with higher peaks` < 0.05)
cases <- cases[,c(1,2,7)]
cases %<>% set_colnames(c('name', 'region', 'residual_q')) 
cases$name.region = paste0(vapply(strsplit(cases$name," "), `[`, 1, FUN.VALUE=character(1)), '.', cases$region)
cases$group <- 'dep_cases'

```

## depudyt controls (controls = patients that survive with at least 5 years follow up)
```{r}

controls <- fread('/mctp/share/users/rebernrj/projects/depuydt_nbl/out/gistic2/controls/all_lesions.conf_90.txt', data.table = F) %>%
  dplyr::filter(`Amplitude Threshold` != 'Actual Copy Change Given') %>%
  dplyr::filter(`Residual q values after removing segments shared with higher peaks` < 0.05)
controls <- controls[,c(1,2,7)]
controls %<>% set_colnames(c('name', 'region', 'residual_q')) 
controls$name.region = paste0(vapply(strsplit(controls$name," "), `[`, 1, FUN.VALUE=character(1)), '.', controls$region)
controls$group <- 'dep_controls'

```

## depudyt all (including other)
case = patients that die within 1.5 years
controls = patients that survive with at least 5 years follow up
other = not meeting either criterion
https://figshare.com/collections/Meta-mining_of_copy_number_profiles_of_high-risk_neuroblastoma_tumors/4169864
```{r}

depudyt_all <- fread('/mctp/share/users/rebernrj/projects/depuydt_nbl/out/gistic2/all/all_lesions.conf_90.txt', data.table = F) %>%
  dplyr::filter(`Amplitude Threshold` != 'Actual Copy Change Given') %>%
  dplyr::filter(`Residual q values after removing segments shared with higher peaks` < 0.05)
depudyt_all <- depudyt_all[,c(1,2,7)]
depudyt_all %<>% set_colnames(c('name', 'region', 'residual_q')) 
depudyt_all$name.region = paste0(vapply(strsplit(depudyt_all$name," "), `[`, 1, FUN.VALUE=character(1)), '.', depudyt_all$region)
depudyt_all$group <- 'dep_all'

```



# Look for regions in common b/t cohorts

## Regions common to all groups 
```{r}

merged <- rbind(cases,controls,depudyt_all,mycn0, mycn1, ours) %>%
  dplyr::mutate(value = 1) %>%
  dplyr::select(-name, -residual_q, -region) %>%
  unique() %>%
  spread(., group, value = value, fill =0) %>%
  rowwise() %>%
  dplyr::mutate(total = sum(c_across(dep_all:our_nbl_cohort)))


```



```{r}

temp <- fread('/mctp/share/users/rebernrj/projects/nbl/data/clinical/nbl.clinical.cogmeet.20210525.tsv') %>%
  dplyr::filter(uf.cured == 'yes')

```

