---
title: "prep_depuydt_files_gistic2"
author: "Ryan Rebernick"
date: "6/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
```{r}

#install.packages(c('dplyr', 'magrittr', 'data.table'))

library(data.table)
library(dplyr)
library(magrittr)

```



# Prepare segmentation file
Instead of chrX/Y labels as chr23. For all samples the chr23 shows up as having 0 markers. This chromosome is excluded from the segment file.
```{r}

seg <- fread('/mctp/share/users/rebernrj/projects/depuydt_nbl/data/figshare/table3_segmented_copy_number_data.txt') %>%
  mutate(numMarkers = -1)

# loop through matrix files 
dir <- '/mctp/share/users/rebernrj/projects/depuydt_nbl/data/GEO/GSE103123/GSE103123_processed_matrix_tables/'
files <- list.files(path = dir, full.names = T)

for(f in files){
  # get current matrix and samples contained in the matrix
  print(f)
  mat <- fread(f)
  ## ensure proper formatting of chr
  mat$chr <- gsub('chr', '', mat$chr)
  ## remove values that do not correspond to sample numbers
  samples <- as.numeric(colnames(mat))
  samples <- as.character(samples[!is.na(samples)])
  
  for(s in samples){
    # hold for current segment
    curSeg <- seg[Name ==s,]
    for(x in 1:length(curSeg$Name)){
      # assign number of markers
      curMat <- mat %>% dplyr::filter(chr == curSeg$chromosome[x], position <= curSeg$max[x], position >= curSeg$min[x])
      curSeg$numMarkers[x] <- length(curMat$position)
    }
    # reassign to segment file
    seg[Name ==s] <- curSeg
  }
  
}

# Exclude chr23
seg %<>% 
  dplyr::select(Name, chromosome, min, max, numMarkers, logratio) %>%
  dplyr::filter(chromosome != 'chr23')
colnames(seg) <- c('sample', 'chromosome', 'start', 'end', 'numMarkers', 'logRatio')

# save 
write.table(seg, paste0('/mctp/share/users/rebernrj/projects/depuydt_nbl/int/gistic2/depuydt_seg_', format(Sys.time(), "%Y-%m-%d"), '.txt'), row.names = F, col.names = T, sep = '\t', quote = F)




```



# Prepare markers file
Very little overlap between files. Use pseudo markers through GISTIC
```{r}

dir <- '/mctp/share/users/rebernrj/projects/depuydt_nbl/data/GEO/GSE103123/GSE103123_processed_matrix_tables/'
files <- list.files(path = dir, full.names = T)

mat <- fread(files[1]) %>% 
  dplyr::select(chr, position) %>%
  dplyr::mutate(chr = gsub('chr', '', chr)) %>%
  dplyr::mutate(chr.pos = paste0(chr, '.', position)) %>%
  filter(chr %in% 1:22)
mat2 <- fread(files[2]) %>% 
  dplyr::select(chr, position) %>% 
  dplyr::mutate(chr = gsub('chr', '', chr)) %>%
  dplyr::mutate(chr.pos = paste0(chr, '.', position)) %>%
  filter(chr %in% 1:22)
chr.pos <- mat$chr.pos[mat$chr.pos %in% mat2$chr.pos]

for(x in 3:length(files)){
  print(paste("X:", x))
  curMat <- fread(files[x]) %>% 
  dplyr::mutate(chr = gsub('chr', '', chr)) %>%
  dplyr::select(chr, position) %>% 
  dplyr::mutate(chr.pos = paste0(chr, '.', position)) %>%
  filter(chr %in% 1:22)
  chr.pos <- chr.pos[chr.pos %in% curMat$chr.pos]
  print(paste('length of chr.pos:', length(chr.pos)))
}


```



# Prepare array list file
```{r}

# all samples
f <- fread('/mctp/share/users/rebernrj/projects/depuydt_nbl/data/figshare/table1_sample_metadata.txt') %>%
  dplyr::select(Number) %>%
  set_colnames('array')
#write.table(f, paste0('/mctp/share/users/rebernrj/projects/depuydt_nbl/int/gistic2/depuydt_array_list_all_', format(Sys.time(), "%Y-%m-%d"), '.txt'), row.names = F, col.names = T, sep = '\t', quote = F)


# 'Cases' - those that die w/in 1.5 years
f <- fread('/mctp/share/users/rebernrj/projects/depuydt_nbl/data/figshare/table1_sample_metadata.txt') %>%
  dplyr::filter(Class == 'case') %>%
  dplyr::select(Number) %>%
  set_colnames('array')
#write.table(f, paste0('/mctp/share/users/rebernrj/projects/depuydt_nbl/int/gistic2/depuydt_array_list_cases_', format(Sys.time(), "%Y-%m-%d"), '.txt'), row.names = F, col.names = T, sep = '\t', quote = F)

# 'Controls' - those that survive at least 5 years follow up
f <- fread('/mctp/share/users/rebernrj/projects/depuydt_nbl/data/figshare/table1_sample_metadata.txt') %>%
  dplyr::filter(Class == 'control') %>%
  dplyr::select(Number) %>%
  set_colnames('array')
#write.table(f, paste0('/mctp/share/users/rebernrj/projects/depuydt_nbl/int/gistic2/depuydt_array_list_controls_', format(Sys.time(), "%Y-%m-%d"), '.txt'), row.names = F, col.names = T, sep = '\t', quote = F)

# Non-MYCN amplified
f <- fread('/mctp/share/users/rebernrj/projects/depuydt_nbl/data/figshare/table1_sample_metadata.txt') %>%
  dplyr::filter(MYCN == 0) %>%
  dplyr::select(Number) %>%
  set_colnames('array')
write.table(f, paste0('/mctp/share/users/rebernrj/projects/depuydt_nbl/int/gistic2/depuydt_array_list_MYCN0_', format(Sys.time(), "%Y-%m-%d"), '.txt'), row.names = F, col.names = T, sep = '\t', quote = F)

# MYCN amplified
f <- fread('/mctp/share/users/rebernrj/projects/depuydt_nbl/data/figshare/table1_sample_metadata.txt') %>%
  dplyr::filter(MYCN == 1) %>%
  dplyr::select(Number) %>%
  set_colnames('array')
write.table(f, paste0('/mctp/share/users/rebernrj/projects/depuydt_nbl/int/gistic2/depuydt_array_list_MYCN1_', format(Sys.time(), "%Y-%m-%d"), '.txt'), row.names = F, col.names = T, sep = '\t', quote = F)

```



# Prepare CNV file
This file is for specifying germ line CNVs to be excluded from the analysis. This information is not available in the manuscript this data derives from. A blank file is created.
```{r}

# empty df with appropriate colnames
f <- data.frame(matrix(ncol = 6, nrow = 0))
x <- c('cnv_identifier', 'chr', 'narrow_start', 'narrow_end', 'wide_start', 'wide_end')
colnames(f) <- x

write.table(f, paste0('/mctp/share/users/rebernrj/projects/depuydt_nbl/int/gistic2/depuydt_cnv_', format(Sys.time(), "%Y-%m-%d"), '.txt'), row.names = F, col.names = T, sep = '\t', quote = F)

```

